---
title: "R Notebook"
output:
  html_notebook: default
  html_document: default
---
# Spearman Correlations
```{r echo = FALSE, message = FALSE}
library(knitr)
# Prepare the watershed characteristic data.
source("prep/prep_metrics.R")
#==============================================================================
# Create a training data set and testing data set.
#==============================================================================
base.cur <- train_test(base.cur)
base.imp <- train_test(base.imp)
#==============================================================================
# Alteration columns.
alt.cols <- c("ALTERATION_MH21", "ALTERATION_DH17",
              "ALTERATION_HIGH_PULSE_COUNT", "ALTERATION_FLASHINESS",
              "ALTERATION_LOW_PULSE_DURATION", "ALTERATION_3_DAY_MAXIMUM",
              "ALTERATION_3_DAY_MINIMUM", "ALTERATION_EXTREME_LOW_FLOW_FREQ",
              "ALTERATION_EXTREME_LOW_FLOW_DURATION", "IMPERV")
alt.cols <- c("ALTERATION_MH21", "ALTERATION_FLASHINESS",
              "ALTERATION_LOW_PULSE_DURATION", "IMPERV")
# Import correlation functions.
source("functions/correlation_func.R")
```
### Watershed Characteristics
```{r echo = FALSE}
kable(corr_env_env(base.cur, "cur"),
             caption = "Watershed Chracteristics", row.names = FALSE,
      align = "lccccccccc")
```
### Baseline/Current Scenario Alteration x Alteration
```{r echo = FALSE}
knitr::kable(corr_metric_metric(base.cur, "cur"), row.names = FALSE,
      align = "lccccccccc")
```
### Baseline/Impervious Scenario Alteration x Alteration
```{r echo = FALSE}
knitr::kable(corr_metric_metric(base.imp, "imp"), row.names = FALSE,
      align = "lccccccccc")
```
### Baseline/Current Scenario Alteration x Watershed Characteristic
```{r echo = FALSE}
knitr::kable(corr_metric_env(base.cur, "cur"), row.names = FALSE,
      align = "lcccccccc")
```
### Baseline/Impervious Scenario Alteration x Watershed Characteristic
```{r echo = FALSE}
knitr::kable(corr_metric_env(base.imp, "imp"), row.names = FALSE,
      align = "lcccccccc")
```
### Baseline/Current Scenario Alteration x %Impervious Cover
```{r echo = FALSE}
knitr::kable(corr_metric_imp(base.cur, "cur"), row.names = FALSE,
      align = "lc")
```
### Baseline/Impervious Scenario Alteration x %Impervious Cover
```{r echo = FALSE}
knitr::kable(corr_metric_imp(base.imp, "imp"), row.names = FALSE,
      align = "lc")
```
***
# Loess Curves

The color gradient in all loess plots belwo is based on the percentage of impervious
cover in the Current scenario. The percentage of impervious cover in each
watershed from the Baseline scenario is approximately zero. To create the 
Impervious Cover scenario, the percentage of impervious cover in each watershed
was increased by 10%.


### Baseline/Current Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}
#==============================================================================
#setwd("C:/Users/zsmith/Desktop/Impervious_Cover/Output/loess")
source("functions/loess_func.R")
#==============================================================================
metric.vec <- c("ALTERATION_FLASHINESS", "ALTERATION_LOW_PULSE_DURATION")
char.vec <- c("AREA", "KARST", "PRECIP", "FCODE", "SLOPE", "SOIL")
#-----------------------------------------------------------------------------
flash <- lapply(char.vec, function(char.x) {
  plot_loess(base.cur, "ALTERATION_FLASHINESS", watershed.char =  char.x,
             low.lim = 0, high.lim = 10)
})
#png("ALTERATION_FLASHINESS_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(flash[[1]], flash[[2]], flash[[3]], flash[[4]], flash[[5]], flash[[6]])
#dev.off()
#==============================================================================


```
### Baseline/Impervious Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}
flash <- lapply(char.vec, function(char.x) {
  plot_loess(base.imp, "ALTERATION_FLASHINESS", watershed.char =  char.x,
             low.lim = 0, high.lim = 10)
})
#png("ALTERATION_FLASHINESS_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(flash[[1]], flash[[2]], flash[[3]], flash[[4]], flash[[5]], flash[[6]])
#dev.off()
```
### Baseline/Current Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}
low.pulse <- lapply(char.vec, function(char.x) {
  
  plot_loess(base.cur, "ALTERATION_LOW_PULSE_DURATION", watershed.char =  char.x,
             low.lim = 0, high.lim = 10)
  
})
#png("ALTERATION_LOW_PULSE_DURATION_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(low.pulse[[1]], low.pulse[[2]], low.pulse[[3]],
             low.pulse[[4]], low.pulse[[5]], low.pulse[[6]],
             ncol = 2, nrow = 3)
#dev.off()
```

### Baseline/Impervious Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}

#==============================================================================
low.pulse <- lapply(char.vec, function(char.x) {
  
  plot_loess(base.imp, "ALTERATION_LOW_PULSE_DURATION", watershed.char =  char.x,
             low.lim = 0, high.lim = 10)
  
})
#png("ALTERATION_LOW_PULSE_DURATION_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(low.pulse[[1]], low.pulse[[2]], low.pulse[[3]],
             low.pulse[[4]], low.pulse[[5]], low.pulse[[6]],
             ncol = 2, nrow = 3)
#dev.off()
```
*** 
# Random Forest
### Baseline/Current Scenario
```{r echo = FALSE, message = FALSE}
source("functions/random_forest_func.R")
plot_random_forest(base.cur[base.cur$TYPE %in% "train", ], "cur", alt.cols, 100000)
```
### Baseline/Impervious Scenario
```{r echo = FALSE, message = FALSE}
plot_random_forest(base.imp[base.imp$TYPE %in% "train", ], "imp", alt.cols, 100000)
```
***
# Rpart
### Baseline/Current Scenario
```{r echo = FALSE}
source("functions/rpart_func.R")
base.cur %>% 
  filter(TYPE %in% "train") %>% 
  plot_rpart("cur", alt.cols, "All")
```
***
### Baseline/Impervious Scenario
```{r echo = FALSE}
base.imp %>% 
  filter(TYPE %in% "train") %>% 
  plot_rpart("imp", alt.cols, "All")
```







