---
title: "R Notebook"
output:
  html_notebook: default
  html_document: default
  pdf_document: default
  word_document: default
---

# Spearman Correlations
```{r echo = FALSE, message = FALSE}
library(knitr)
library(dplyr)
#library(formattable)
# Prepare the watershed characteristic data.
source("prep/prep_metrics.R")
#==============================================================================
# Create a training data set and testing data set.
#==============================================================================
base.cur <- train_test(base.cur)
base.imp <- train_test(base.imp)
#==============================================================================
# Alteration columns.
alt.cols <- c("ALTERATION_MH21", "ALTERATION_DH17",
              "ALTERATION_HIGH_PULSE_COUNT", "ALTERATION_FLASHINESS",
              "ALTERATION_LOW_PULSE_DURATION", "ALTERATION_3_DAY_MAXIMUM",
              "ALTERATION_3_DAY_MINIMUM", "ALTERATION_EXTREME_LOW_FLOW_FREQ",
              "ALTERATION_EXTREME_LOW_FLOW_DURATION", "IMPERV")
env.cols <- c("AREA", "KARST", "PRECIP", "KFACT", "AVGKW", "FCODE", 
               "SLOPE", "SOIL", "ECOREGION_C", "ECOREGION4", "ECOREGION3")

# Import correlation functions.
source("functions/correlation_func.R")
```

### Watershed Characteristics
Soil Erodibility Factor and Mean Soil Erodibility were significantly correlated
with the other watershed characteristics (r > 0.5), and therefore, were 
removed from the analysis.

```{r echo = FALSE}
kable(spear_corr(base.cur, "cur", env.cols, type = "env_env"),
             caption = "Watershed Chracteristics", row.names = FALSE,
      align = "lccccccccc")
```

```{r echo = FALSE}
#env.cols <- c("AREA", "KARST", "PRECIP", "FCODE", "SLOPE", "SOIL")
```

##  Alteration x Alteration
In both scenarios 3-Day Minimum (cfs/sqmi), Extreme Low Flow Frequency,
and Extreme Low Flow Duration were ignored because they were not present in the
Phase 1 analysis or the Middle Potomac study. In subsequent analyses, these
three alteration metrics were removed.

In the Baseline/Current scenario Flashiness was significantly correlated 
(r > 0.50) with MH21, DH17, High Pulse Count, and 3-Day Maximum (cfs/sqmi).
However, in the Baseline/Impervious scenario Flashiness was only significantly
correlated with High Pulse Count and 3-Day Maximum (cfs/sqmi). DH17 and MH21 
were significantly correlated (r = 0.86). MH21 was excluded becuase it was 
slightly more correlated with Flashiness (r = -0.40). Flashiness and 
DH17 were retained for further analysis.

In both scenarios, Low Pulse Duration was not significantly correlated with
any of the other alteration metrics. Therefore, Low Pulse Duration was retained
for further analyses.

### Baseline/Current Scenario


```{r echo = FALSE}
knitr::kable(corr_metric_metric(base.cur, "cur", alt.cols), row.names = FALSE,
      align = "lccccccccc")
```

### Baseline/Impervious Scenario
```{r echo = FALSE}
knitr::kable(corr_metric_metric(base.imp, "imp", alt.cols), row.names = FALSE,
      align = "lccccccccc")
```

```{r echo = FALSE}
alt.cols <- alt.cols[!alt.cols %in% c("ALTERATION_3_DAY_MINIMUM",
                                         "ALTERATION_EXTREME_LOW_FLOW_FREQ",
                                         "ALTERATION_EXTREME_LOW_FLOW_DURATION")]
```


## Alteration x Watershed Characteristic
In both scenarios Low Pulse Duration was not significantly correlated
(-0.5 < r < 0.5) with any watershed characteristics.

In the Baseline/Current scenario, DH17 was significantly correlated with
Soil Erodibility Factor, Mean Soil Erodbility. However, DH17 was
not signficantly correlated with any watershed characteristics in
the Baseline/Impervious scenario. 

In the Baseline/Current scenario, Flashiness was significantly correlated with
FCODEgraphic Province and Average Watershed Slope. However, Flashiness was
only significanly correlated with Average watershed Slop in the 
Baseline/Impervious scenario. In both scenarios Flashiness was negatively 
correlated with Average Watershed Slope but the relationship was not as strong
in the Baseline/Impervious scenario.

### Baseline/Current Scenario 
```{r echo = FALSE}
knitr::kable(corr_metric_env(base.cur, "cur", env.cols, alt.cols), row.names = FALSE,
      align = "lcccccccc")
```

### Baseline/Impervious Scenario
```{r echo = FALSE}
knitr::kable(corr_metric_env(base.imp, "imp", env.cols, alt.cols), row.names = FALSE,
      align = "lcccccccc")
```
## Alteration x %Impervious Cover
The percent change in Impervious Cover was held constant in the 
Baseline/Impervious scenario, and therefore, correlations will not 
exist between the Alteration metrics and the percent change in 
Impervious cover. Low Pulse Duration was not signficantly correlated with
the Percentage of Impervious Cover Change in the Baseline/Current Scenario. 
Percent Impervious Cover Change was negatively correlated with DH17 and 
positively correlated with Flashiness in the Baseline/Current scenario. 

### Baseline/Current Scenario 
```{r echo = FALSE}
knitr::kable(corr_metric_imp(base.cur, "cur", alt.cols), row.names = FALSE,
      align = "lc")
```

---

# Loess Curves
The color gradient in all loess plots belwo is based on the percentage of impervious
cover in the Current scenario. The percentage of impervious cover in each
watershed from the Baseline scenario is approximately zero. To create the 
Impervious Cover scenario, the percentage of impervious cover in each watershed
was increased by 10%.

Categorical variables, like FCODE (FCODEgraphic Province) and Soil, are not
appropriate for Loess curves. However, I left these plots in because they
have the potental to inform patterns in the other plots.

In general, the Baseline/Impervious Loess Curves show similar but weaker trends
to the Baseline/Current. 

## Flashiness

Flashiness in the Baseline/Current scenario had a slight concave unimodal
distribution when plotted against Area. This pattern did not exist for
the Baseline/Impervious scenario.  Most likely the pattern observed in the
Baseline/Current had to do with a development bias. The percent change
in impervious cover can change more rapidly in small areas.  For example,
a equally sized parking lots could be added to a small watershed and a 
large watershed. Relative to the the large watershed, the proportion of 
area disturbed by the parking lot is greater in the smaller watershed. 
Therfore, minor land cover changes that occur in small Baseline/Current
watersheds may have a large influence on Flashiness values. When land cover
changes are held constant in the Baseline/Current scenario, the concave
unimodal distribution disappears.

The relationship beteen Karst and Flashiness appears to be 
slightly stronger in the Baseline/Impervious scenario. As the percentage of
Karst in a watershed increases, Flashiness may also increase.

In both scenarios, Percipitation had an interesting convex unimodal 
distribution when plotted against Flashiness. Although Precipitation was not 
sigificantly correlated with FCODEgraphic Province (r = -0.45), the most 
obvious explanation for this unimodal distiribution would be that certain 
areas receive more rain than others. One way to look into this further would
be to dividethe Ridge and Valley FCODEgrapic Province (FCODE = 62) into two 
seperate provinces. My current assumption is that Ridges receive more rain than
the other provinces but this is being masked by grouping Ridges with Valleys.

Slope was negatively correlated with Flashiness in both scenarios 
(See correlation section). Development is more prominent in flat 
areas (low slope), therefore, in the Baseline/Current scenario the negative 
correlation was assumed to be an artifact development bias. Although the 
Baseline/Imperious scenario correlation was weaker, there does appear to be a
negative correlation between Slope and Flashiness, even if the percent change
in impervious cover is a constant.

### Baseline/Current Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}
#==============================================================================
#setwd("C:/Users/zsmith/Desktop/Impervious_Cover/Output/loess")
source("functions/loess_func.R")
#==============================================================================
metric.vec <- c("ALTERATION_FLASHINESS", "ALTERATION_LOW_PULSE_DURATION")
char.vec <- c("AREA", "KARST", "PRECIP", "FCODE", "SLOPE", "SOIL")
#***************************************************************************--
flash <- lapply(char.vec, function(char.x) {
  plot_loess(base.cur, "ALTERATION_FLASHINESS", watershed.char =  char.x,
             low.lim = 0, high.lim = 10, color = TRUE)
})
#png("ALTERATION_FLASHINESS_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(flash[[1]], flash[[2]], flash[[3]], flash[[4]], flash[[5]], flash[[6]])
#dev.off()
#==============================================================================
```

### Baseline/Impervious Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}
flash <- lapply(char.vec, function(char.x) {
  plot_loess(base.imp, "ALTERATION_FLASHINESS", watershed.char =  char.x,
             low.lim = 0, high.lim = 10)
})
#png("ALTERATION_FLASHINESS_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(flash[[1]], flash[[2]], flash[[3]], flash[[4]], flash[[5]], flash[[6]])
#dev.off()
```
## DH17
Area and Karst had similar distirbutions in both scenarios but the values were
generally lower and variability was generally smaller in the 
Baseline/Impervious scenario.

The Baseline/Current scenario had a concave, almost, biomodal distribution 
between Precipitation and DH17. This pattern was dramatically reduced in 
the Baseline/Impervious scenario.  There did appear to be a positive 
relationship between Perceiptation and DH17 for percipitation values
greater than 42.

The was a strong positive correlation between Slope and DH17 in the 
Baseline/Current scenario. However, this relationship did not exist in the
Baseline/Impervious scenario. As was stated in the Flashiness section, the
relationship observed in the Baseline/Current scenario may be due to a
development bias.  Development occurs more frequently in flat areas.

### Baseline/Current Scenario

```{r echo = FALSE, warning = FALSE, message = FALSE}
dh17 <- lapply(char.vec, function(char.x) {
  
  plot_loess(base.cur, "ALTERATION_DH17", watershed.char =  char.x,
             low.lim = 0, high.lim = 10, color = TRUE)
  
})
#png("ALTERATION_LOW_PULSE_DURATION_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(dh17[[1]], dh17[[2]], dh17[[3]],
             dh17[[4]], dh17[[5]], dh17[[6]],
             ncol = 2, nrow = 3)
#dev.off()
```

### Baseline/Impervious Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}

#==============================================================================
dh17 <- lapply(char.vec, function(char.x) {
  
  plot_loess(base.imp, "ALTERATION_DH17", watershed.char =  char.x,
             low.lim = 0, high.lim = 10)
  
})
#png("ALTERATION_LOW_PULSE_DURATION_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(dh17[[1]], dh17[[2]], dh17[[3]],
             dh17[[4]], dh17[[5]], dh17[[6]],
             ncol = 2, nrow = 3)
#dev.off()
```
## Low Pulse Duration

Area and Karst had similar distirbutions in both scenarios but the values were
generally lower and variability was generally smaller in the 
Baseline/Impervious scenario.

In the Baseline/Current scenario, the relationship between Precipiation
and Low Pulse Duration was a slightly concave unimodal distribution. There
was no relationship in the Basline/Impervious scenario and there was
generally less variance.

There was a positive relationship between Slope and Low Pulse Duration
in the Baseline/Current scenario.  There was no relationship observed
between Slope and Low Pulse Duration in the Basline/Impervious scenario.

### Baseline/Current Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}
low.pulse <- lapply(char.vec, function(char.x) {
  
  plot_loess(base.cur, "ALTERATION_LOW_PULSE_DURATION", watershed.char =  char.x,
             low.lim = 0, high.lim = 10, color = TRUE)
  
})
#png("ALTERATION_LOW_PULSE_DURATION_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(low.pulse[[1]], low.pulse[[2]], low.pulse[[3]],
             low.pulse[[4]], low.pulse[[5]], low.pulse[[6]],
             ncol = 2, nrow = 3)
#dev.off()
```

### Baseline/Impervious Scenario
```{r echo = FALSE, warning = FALSE, message = FALSE}

#==============================================================================
low.pulse <- lapply(char.vec, function(char.x) {
  
  plot_loess(base.imp, "ALTERATION_LOW_PULSE_DURATION", watershed.char =  char.x,
             low.lim = 0, high.lim = 10)
  
})
#png("ALTERATION_LOW_PULSE_DURATION_loess_plots.png", width = 6.5, height = 10.5, units = "in", res = 800)
grid.arrange(low.pulse[[1]], low.pulse[[2]], low.pulse[[3]],
             low.pulse[[4]], low.pulse[[5]], low.pulse[[6]],
             ncol = 2, nrow = 3)
#dev.off()
```

---

# Random Forest
In the Baseline/Current scenario, Slope was consistently the most important 
watershed characteristic for predicting Alteration Metric values.  Precipiation
and Area were also important in the Flashiness Random Forest model. Similaraly,
Precipiation was the second most import preditive variable in the Low Pulse
Duration Random Forest model.

In the Baseline/Impervious scenario, Slope dropped out as the most important
predictive variable and was replaced with Precipiation. Karst was the second most
important predictive variable for the DH17 Random Forest model. For the 
Flashiness Random Forest model, Slope and Karst were also important predictive
variables. Slope was the second most important predictive variable for the
Low Pulse Duration Random Forest model but appeared to be relatively weak.

### Baseline/Current Scenario
```{r echo = FALSE, message = FALSE}
source("functions/random_forest_func.R")
base.cur %>% 
  #filter(TYPE %in% "train") %>% 
  plot_random_forest("cur", alt.cols, 100000)
```

### Baseline/Impervious Scenario
```{r echo = FALSE, message = FALSE}
base.imp %>% 
  #filter(TYPE %in% "train") %>%
  plot_random_forest("imp", alt.cols, 100000)
```

# Rpart

The _caret_ package function `train` was used to create the rpart trees. 
A `maxdepth` of 4 was used. Addtionally, `trainControl` was set to 
method = "repeatedcv", number = 10, repeats = 100; in other words a 10 fold
cross validation was performed 100 times.

### Baseline/Current Scenario
```{r echo = FALSE}
source("functions/rpart_func.R")
base.cur %>% 
  filter(TYPE %in% "train") %>% 
  plot_rpart("cur", alt.cols, "All") %>% 
  invisible()
```

### Baseline/Impervious Scenario
```{r echo = FALSE}
base.imp %>% 
  filter(TYPE %in% "train") %>% 
  plot_rpart("imp", alt.cols, "All") %>% 
  invisible()
```







